Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Compound figure
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@16 db7c848b-7755-4df2-b763-8796797afa52

@@

Sequal is a simplification layer for Equalizer. It is based on the realization
that while fully expressive, the verbatim abstraction layer of nodes, pipes,
windows and channels in Equalizer requires significant learning to fully
understand and exploit. In reality, a higher abstraction of \textsf{Application}
and \textsf{Renderer} is sufficient for most use cases. In Sequel, the
application class drives the configuration, and one renderer instance is created
for each (pipe) render thread. They also provide the natural place to store and
distribute data. Last, but not least, view data provides a convenient way to
manage multiple views by storing the camera, model or any other view-specific
information.

@@

The initial version of Equalizer implemented sort-first (2D), sort-last (DB) and
stereo (EYE) decomposition \cite{EMP:09}. In the following we present new
decomposition modes and motivate their use case.

Time-multiplexing, or DPlex, was already implemented in \cite{BRE:05}. While it
increases the framerate linearly, it does not decrease the latency between user
input and the corresponding output. Consequently, this decomposition mode is
mostly useful for non-interactive movie generation. It is transparent to
Equalizer applications, but does require the configuration latency to be equal
or greater than the number of source channels. Furthermore, to work in a
multi-threaded, multi-GPU configuration, the application needs to support
running the rendering threads asynchronously, as outlined in
\sref{SEC_threading}. The output frame rate of the destination channel is
smoothened using a frame rate equalizer (\sref{SEC_framerateEq}).

Tile and chunk decompositions are a variant of sort-first and sort-last
rendering, respectively. They decompose the scene into a predefined set of
fixed-size image tiles or database ranges. These tasks are queued and processed
by all source channels by polling a server-central queue. Prefetching ensures
that the task communication overlaps with rendering. As shown in \cite{SPEP:16},
these modes provide better performance due to being inherently load-balanced, as
long as there is an insignificant overhead for the render task setup. This mode
is transparent to Equalizer applications.

Pixel compounds decompose the destination channel by interleaving rows or
columns in image space. They are a variant of sort-first decomposition which
works well for fill-limited applications which are not geometry bound. Source
channels cannot reduce geometry load through view frustum culling, since each
source channel has almost the same frustum (only shifted by some pixels),
but applied to a reduced 2D viewport. However, the fragment load on all source
channels is very similar due to the interleaved distribution of pixels. This
functionality is transparent to Equalizer applications, and the default
compositing implementation uses the OpenGL stencil buffer to blit pixels onto
the destination channel.

Subpixel compounds are similar to pixel compounds, but they decompose the work
for a single pixel, for example when using multisampling or depth of
field. Composition typically uses accumulation and averaging of all computed
fragments for a pixel. This feature is not fully transparent to the application,
since it needs to adapt (jitter or tilt) the frustum based on the iteration
executed. Furthermore, subpixel compounds interact with idle image refinements,
e.g., they can accelerate idle anti-aliasing of a scene when the camera and
scene are not changed.

Equalizer are an addition to compound trees. They modify parameters of their
respective subtree at runtime to optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behaviour.

\subsubsection{Sort-First and Sort-Last Load Equalizer}

>>>>

This section introduces new decomposition modes, runtime adaptations and other
optimizations to increase the rendering performance for a wide variety of
algorithms.

The initial version of Equalizer implemented sort-first (2D), sort-last (DB),
stereo (EYE) and multilevel decompositions \cite{EMP:09}. In the following we present new
decomposition modes and motivate their use case. \fig{fig:compounds} provides an
overview of the new modes.

\begin{figure*}[ht]\center
  \subfigure[Time-Multiplex]{\label{fig:dplex}
    \includegraphics[width=0.24\textwidth]{images/dplex}}\hfil
  \subfigure[Tiles]{\label{fig:tile}
    \includegraphics[width=0.24\textwidth]{images/tile}}\hfil
  \subfigure[Pixel]{\label{fig:pixel}
    \includegraphics[width=0.24\textwidth]{images/pixel}}\hfil
  \subfigure[Subpixel]{\label{fig:subpixel}
    \includegraphics[width=0.24\textwidth]{images/subpixel}}\hfil
  \caption{New Equalizer task decompositions for parallel rendering}
  \label{fig:compounds}
\end{figure*}

Time-multiplexing (\fig{fig:dplex}), or DPlex, was implemented in \cite{BRE:05}
for shared memory machines. It is however a better fit for distributed memory
systems, since the separate memory space makes concurrent rendering of different
frames easier to implement. While it increases the framerate linearly, it does
not decrease the latency between user input and the corresponding
output. Consequently, this decomposition mode is mostly useful for
non-interactive movie generation. It is transparent to Equalizer applications,
but does require the configuration latency to be equal or greater than the
number of source channels. Furthermore, to work in a multi-threaded, multi-GPU
configuration, the application needs to support running the rendering threads
asynchronously, as outlined in \sref{SEC_threading}. The output frame rate of
the destination channel is smoothened using a frame rate equalizer
(\sref{SEC_framerateEq}).

Tile (\fig{fig:tile}) and chunk decompositions are a variant of sort-first and
sort-last rendering, respectively. They decompose the scene into a predefined
set of fixed-size image tiles or database ranges. These tasks are queued and
processed by all source channels by polling a server-central queue. Prefetching
ensures that the task communication overlaps with rendering. As shown in
\cite{SPEP:16}, these modes provide better performance due to being inherently
load-balanced, as long as there is an insignificant overhead for the render task
setup. This mode is transparent to Equalizer applications.

Pixel compounds (\fig{fig:pixel}) decompose the destination channel by
interleaving rows or columns in image space. They are a variant of sort-first
decomposition which works well for fill-limited applications which are not
geometry bound. Source channels cannot reduce geometry load through view frustum
culling, since each source channel has almost the same frustum (only shifted by
some pixels), but applied to a reduced 2D viewport. However, the fragment load
on all source channels is very similar due to the interleaved distribution of
pixels. This functionality is transparent to Equalizer applications, and the
default compositing implementation uses the OpenGL stencil buffer to blit pixels
onto the destination channel.

Subpixel compounds (\fig{fig:subpixel}) are similar to pixel compounds, but they
decompose the work for a single pixel, for example when using multisampling or
depth of field. Composition typically uses accumulation and averaging of all
computed fragments for a pixel. This feature is not fully transparent to the
application, since it needs to adapt (jitter or tilt) the frustum based on the
iteration executed. Furthermore, subpixel compounds interact with idle image
refinements, e.g., they can accelerate idle anti-aliasing of a scene when the
camera and scene are not changed.

Equalizer are an addition to compound trees. They modify parameters of their
respective subtree at runtime to optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behaviour.
\subsubsection{Sort-First and Sort-Last Load Equalizer}

@@

Sequal is a simplification layer for Equalizer. It is based on the realization
that while fully expressive, the verbatim abstraction layer of nodes, pipes,
windows and channels in Equalizer requires significant learning to fully
understand and exploit. In reality, a higher abstraction of \textsf{Application}
and \textsf{Renderer} is sufficient for most use cases. In Sequel, the
application class drives the configuration, and one renderer instance is created
for each (pipe) render thread. They also provide the natural place to store and
distribute data. Last, but not least, \textsf{ViewData} provides a convenient
way to manage multiple views by storing the camera, model or any other
view-specific information.

@@
